---
title: "p8105_hw2_yf2563"
author: "Yatong Feng"
date: "9/30/2020"
output: github_document
---

```{r setup, message=FALSE}
#load packages
library(tidyverse)
library(readxl)
```

## Problem 1

- specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
- use reasonable variable names
- omit rows that do not include dumpster-specific data
- round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

First, define a path and read the dataset

```{r}
path_to_data = './data/Trash-Wheel-Collection-Totals-8-6-19.xlsx'
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls, digits = 0),
		sports_balls = as.integer(sports_balls)
	) %>% 
  relocate(dumpster,month,year,date,sports_balls)
trashwheel_df
```

