---
title: "p8105_hw2_yf2563"
author: "Yatong Feng"
date: "9/30/2020"
output: github_document
---

```{r setup, message=FALSE}
#load packages
library(tidyverse)
library(readxl)
```

## Problem 1

- specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
- use reasonable variable names
- omit rows that do not include dumpster-specific data
- round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

First, define a path and read the dataset

```{r}
path_to_data = './data/Trash-Wheel-Collection-Totals-8-6-19.xlsx'
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls, digits = 0),
		sports_balls = as.integer(sports_balls)
	) %>% 
  relocate(dumpster,month,year,date,sports_balls)
trashwheel_df
```

Read and clean precipitation data for 2017 and 2018. 

- omit rows without precipitation data and add a variable year

```{r}
# data for 2018
precip_2018 = 
	read_excel(
		path = path_to_data,
		sheet = "2018 Precipitation",
		range = "A2:B14"
	) %>% 
	janitor::clean_names() %>% 
	drop_na(total) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
```

```{r}
# data for 2017
precip_2017 = 
	read_excel(
		path = path_to_data,
		sheet = "2017 Precipitation",
		range = "A2:B14"
	) %>% 
	janitor::clean_names() %>% 
	drop_na(total) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

- combine precipitation datasets and convert month to a character variable

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)

precip_df =
	left_join(precip_df, month_df, by = "month") %>% 
  select(-month) %>% 
  rename(month = month_name) %>%
  relocate(year, month)

precip_df
```


### Summary of the data:

```{r}
summary(trashwheel_df)
summary(precip_df)
```


The first dataset is about the information from the Mr. Trashwheel trash collector in Baltimore, Maryland:

- After filtering, there are a total of `r nrow(trashwheel_df)` rows and `r ncol(trashwheel_df)` columns. 
- The variables include `r names(trashwheel_df)`

The second dataset is about the month precipitation in 2017 and 2018:

- After filtering, there are a total of `r nrow(precip_df)` rows and `r ncol(precip_df)` columns. 
- The variables include `r names(precip_df)`
- The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
- The median number of sports balls found in a dumpster in 2018 was `r trashwheel_df %>% filter(year == 2018) %>% pull(sports_balls) %>% median()`
- The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.
- The total precipitation in 2017 was `r precip_df %>% filter(year == 2017) %>% pull(total) %>% sum()` inches.

## Problem 2

Read and clean the data:

- retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance
- Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r, message=FALSE}
path_prob2 = './data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv'

subway_df =
  read_csv(path_prob2) %>%
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = recode(entry, "YES" = T, "NO" = F))

subway_df
```


