---
title: "p8105_hw2_yf2563"
author: "Yatong Feng"
date: "9/30/2020"
output: github_document
---

```{r setup, message=FALSE}
#load packages
library(tidyverse)
library(readxl)
```

## Problem 1

- specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
- use reasonable variable names
- omit rows that do not include dumpster-specific data
- round the number of sports balls to the nearest integer and converts the result to an integer variable (using as.integer)

First, define a path and read the dataset

```{r}
path_to_data = './data/Trash-Wheel-Collection-Totals-8-6-19.xlsx'
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls, digits = 0),
		sports_balls = as.integer(sports_balls)
	) %>% 
  relocate(dumpster,month,year,date,sports_balls)
trashwheel_df
```

Read and clean precipitation data for 2017 and 2018. 

- omit rows without precipitation data and add a variable year

```{r}
# data for 2018
precip_2018 = 
	read_excel(
		path = path_to_data,
		sheet = "2018 Precipitation",
		range = "A2:B14"
	) %>% 
	janitor::clean_names() %>% 
	drop_na(total) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
```

```{r}
# data for 2017
precip_2017 = 
	read_excel(
		path = path_to_data,
		sheet = "2017 Precipitation",
		range = "A2:B14"
	) %>% 
	janitor::clean_names() %>% 
	drop_na(total) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

- combine precipitation datasets and convert month to a character variable

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)
precip_df = 
	bind_rows(precip_2018, precip_2017)

precip_df =
	left_join(precip_df, month_df, by = "month") %>% 
  select(-month) %>% 
  rename(month = month_name) %>%
  relocate(year, month)

precip_df
```


### Summary of the data:

```{r}
summary(trashwheel_df)
summary(precip_df)
```


The first dataset is about the information from the Mr. Trashwheel trash collector in Baltimore, Maryland:

- After filtering, there are a total of `r nrow(trashwheel_df)` rows and `r ncol(trashwheel_df)` columns. 
- The variables include `r names(trashwheel_df)`

The second dataset is about the month precipitation in 2017 and 2018:

- After filtering, there are a total of `r nrow(precip_df)` rows and `r ncol(precip_df)` columns. 
- The variables include `r names(precip_df)`
- The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
- The median number of sports balls found in a dumpster in 2018 was `r trashwheel_df %>% filter(year == 2018) %>% pull(sports_balls) %>% median()`
- The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.
- The total precipitation in 2017 was `r precip_df %>% filter(year == 2017) %>% pull(total) %>% sum()` inches.

## Problem 2

Read and clean the data:

- retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance
- Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r, message=FALSE}
path_prob2 = './data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv'

subway_df =
  read_csv(path_prob2) %>%
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = recode(entry, "YES" = T, "NO" = F))

subway_df
```

### Summary for the NYC subway data

The data set contains information related to each entrance and exit for each subway station in NYC:

- The data cleaning procedure include:
  - Read the data through read_csv()
  - Clean the column names through clean_names()
  - Retain the required variables through select()
  - Convert the entry variable from character (YES vs NO) to a logical variable through recode()
- After filtering, there are a total of `r nrow(subway_df)` rows and `r ncol(subway_df)` columns, thus the dimension of the dataset is `r dim(subway_df)`.
- The variables include `r names(subway_df)`
- This data is not tidy since the data type of variables route 1 - 11 contains both chr and dbl



### Answer the following questions using these data:

##### 1. How many distinct stations are there? 

```{r}
question1_df =
  subway_df %>% 
  distinct(line, station_name, .keep_all = T)
```

Thus, there are `r nrow(question1_df)` distinct stations


##### 2. How many stations are ADA compliant?

```{r}
question2_df = 
  question1_df %>% 
  filter(ada == TRUE)
```

Thus, `r nrow(question2_df)` stations are ADA compliant

##### 3. What proportion of station entrances / exits without vending allow entrance?

```{r}
question3_without_vending = 
  subway_df %>% 
  filter(vending == "NO")

question3_without_vending_allow_entrance = 
  subway_df %>% 
  filter(vending == "NO") %>% 
  filter(entry == T)

question3_proportion = nrow(question3_without_vending_allow_entrance) / nrow(question3_without_vending)
```

Thus, the proportion of station entrances / exits without vending allow entrance is `r question3_proportion`


Reformat data so that route number and route name are distinct variables.
```{r}
reformat_subway_df =
  question1_df %>%
  mutate(route8 = as.character(route8),
         route9 = as.character(route9),
         route10 = as.character(route10),
         route11 = as.character(route11)) %>%
  pivot_longer(col = route1:route11,
               names_to = "route_name", 
               values_to = "route_number") %>%
  arrange(route_name, route_number)
reformat_subway_df
```


##### 4. How many distinct stations serve the A train? 
```{r}
question4_df =
  question1_df %>% 
  filter(route1 == "A"|route2 =="A"|route3 == "A"|route4 =="A"|
           route5 == "A"|route6 =="A"|route7 == "A"|route8 =="A"|
           route9 == "A"|route10 =="A"|route11 =="A")
```

Thus, there are `r nrow(question4_df)` distinct stations serve the A train

##### 5.Of the stations that serve the A train, how many are ADA compliant?

```{r}
question5_df = 
  question4_df %>% 
  filter(ada == TRUE)
```

Thus, Of the stations that serve the A train, `r nrow(question5_df)` are ADA compliant.

## Problem 3

First, clean the data in pols-month.csv. 

- Use separate() to break up the variable mon into integer variables year, month, and day
- replace month number with month name
- create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r, message=FALSE}
path_pols_month = "./data/pols-month.csv"

pols_month_df = 
  read_csv(path_pols_month) %>%
  janitor::clean_names() %>%
  separate(mon,c("year","month","day")) %>%
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%
  mutate(month = recode(month, `1` = "Jan", `2` = "Feb", `3` = "Mar", `4` = "Apr", `5` = "May" , `6`= "Jun", `7` = "Jul", `8` = "Aug", `9` = "Sep", `10` = "Oct", `11` = "Nov", `12` = "Dec")) %>%
  mutate(president = ifelse(prez_dem == 1, "dem", "gop")) %>%
  select(-prez_dem,-prez_gop,-day) %>%
  arrange(desc(row_number()))

pols_month_df
```


Second, clean the data in snp.csv using a similar process to the above. 

- For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r, message=FALSE}
path_snp = './data/snp.csv'

snp_df = 
  read_csv(path_snp) %>%
  janitor::clean_names() %>%
  separate(date,c("month","day","year")) %>%
  mutate(
    year = as.integer(year), 
    month = as.integer(month), 
    day = as.integer(day)) %>%
  mutate(month = recode(month, `1` = "Jan", `2` = "Feb", `3` = "Mar", `4` = "Apr", `5` = "May" , `6`= "Jun", `7` = "Jul", `8` = "Aug", `9` = "Sep", `10` = "Oct", `11` = "Nov", `12` = "Dec")) %>%
  select(year,month,close)

snp_df
```

Third, tidy the unemployment data.

- Switching from “wide” to “long” format
- Ensuring that key variables have the same name and take the same values.

```{r, message=FALSE}
path_unemployment = './data/unemployment.csv'

unemployment_df = 
  read_csv(path_unemployment) %>%  
  pivot_longer(Jan:Dec,
               names_to = "month", 
               values_to = "unemployment") %>%
  select(year = Year, everything()) %>%
  arrange(desc(row_number()))

unemployment_df
```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
merge_df = left_join(pols_month_df, snp_df, by = c("year","month"))

merge_df = left_join(merge_df, unemployment_df, by = c("year","month"))

merge_df
```

### Summary for the three datasets


The first dataset is pols_month:

- After filtering, there are a total of `r nrow(pols_month_df)` rows and `r ncol(pols_month_df)` columns, thus the dimension is `r dim(pols_month_df)`.
- The variables include `r names(pols_month_df)`
- In this dataset, 'dem' means democratic party, 'gop' means republican party
- The variables of 'gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem' represents the number of governors or senators from republican or democratic party
- The range of years is `r range(pols_month_df$year)`

The second dataset is snp:

- After filtering, there are a total of `r nrow(snp_df)` rows and `r ncol(snp_df)` columns, thus the dimension is `r dim(snp_df)`.
- The variables include `r names(snp_df)`
- The range of years is `r range(snp_df$year)`

The third dataset is about unemployment:

- After filtering, there are a total of `r nrow(unemployment_df)` rows and `r ncol(unemployment_df)` columns, thus the dimension is `r dim(unemployment_df)`.
- The variables include `r names(unemployment_df)`
- The range of years is `r range(unemployment_df$year)`

The resulting dataset is the merge of all of the three datasets:


- After filtering, there are a total of `r nrow(merge_df)` rows and `r ncol(merge_df)` columns, thus the dimension is `r dim(merge_df)`.
- The variables include `r names(merge_df)`
- The range of years is `r range(merge_df$year)`



